Spatial Analysis

### 2021-08-13 Scripts/spatial_analisis.Rmd
Results for both approaches

---
title: "Spatial Analysis"
author: "Juliano Palacios Abrantes"
output:
  # html_document:
  html_notebook:
    fig_width: 6
    fig_height: 4
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float:
      collapsed: false
    theme: darkly
---

@@

# Get state order from North to south
state_order <- my_path("R","Partial/", name = "grid_eez_fp_df.csv",read = T) %>%
  group_by(state) %>% 
  summarise(
    order = mean(lat)
  ) %>% 
  mutate(abrev = c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")
  ) %>% 
  arrange(desc(order)) %>% 
  mutate(order= c("b","a",letters[3:11])) # weirdly maine goes below

@@

# Plotting 
# Make sure st_simplify() is run for eez sf

@@

p_sw <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_eez_sw_sf) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color =abrev , fill = abrev), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    scale_color_manual(values = state_pallet) +
    scale_fill_manual(values = state_pallet) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45))+
    labs(x = "", y = "", title = "A) State waters extrapolation") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot(grid_eez_sw_sf) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    geom_sf(aes(color = abrev),size = 0.1, alpha = 0.3) +
    facet_wrap(~ group) +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet,
                       # labels = unique(grid_eez_sw_sf$group)) +
                       labels = grid_eez_sw_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

@@

p_fp <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_eez_fp_sf) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color = abrev, fill = abrev), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    scale_color_manual(values = state_pallet) +
    scale_fill_manual(values = state_pallet) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45)) +
    labs(x = "", y = "", title = "B) Fishing ports extrapolation") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot(grid_eez_fp_sf) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    geom_sf(aes(color = abrev),size = 0.1, alpha = 0.3) +
    facet_wrap(~ group) +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet,
                       # labels = unique(grid_eez_fp_sf$group)) +
                       labels = grid_eez_fp_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

@@

ggsave(filename = "buffer_figure_b.jpg",
       plot = combined_plot,
       path = my_path("R","Figures"),
       width = 11,
       height = 11
)

@@

# # single species run
# run_tis(input_data = ocean_data,
#         grid = grid_eez_fp_df,
#         years = years,
#         reg = regions,
#         taxa = "Illex illecebrosus",
#         method = "fp"
# )

@@

@@

spp_paths <- list.files(my_path("R",extra_path = "Partial/Interpolation"))

@@

taxon_read <- paste0(my_path("R",extra_path = "Partial/Interpolation"),taxon_read)

# Load all sp in one table
historic_spp_sw <- bind_rows(lapply(taxon_read, fread)) %>% 
  mutate(state = str_to_sentence(state))

unique(historic_spp$spp)

@@

 data_grid <- historic_spp %>% 
  group_by(state,lat,lon,region) %>% 
  summarise(sum = sum(value,na.rm= T), .groups = "drop") %>% # sum of all species
  group_by(state,lat,lon,region) %>% 
  summarise(mean = mean(sum,na.rm= T), .groups = "drop") # average of all years

ggplot(data = subset(data_grid, !is.na(mean))) +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = mean,
      col = mean
    )
  ) +
  geom_sf(data = land_sf, aes()) +
  facet_wrap(~ state  +  region,
             ncol = 4) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_viridis("Mean Interpolation", alpha = 0.8) +
  scale_color_viridis("Mean Interpolation", alpha = 0.8) +
  theme_bw()
```

@@

state_fit <- historic_spp %>% 
  group_by(state,year,region) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  left_join(periods,
            by = "year") %>% 
  group_by(state,order,label,region) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop")

@@

p <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  ggplot() +
  # geom_sf(aes(fill = mean_per)) +
  geom_sf(aes(fill = mean_per, text = paste(state, mean_per,"% of proportion"))) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_wrap(~ region + label, 
             nrow = 2) +
  labs(x = "Latitude", 
       y = "Longitude") +
  my_ggtheme_p(facet_tx_s = 12,
               leg_pos = "bottom") +
  theme(legend.key.width = unit(1,"line")) +
  theme_dark()

@@

state_fit <- historic_spp %>% 
  group_by(state,year,region,spp) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  left_join(periods,
            by = "year") %>% 
  group_by(state,order,label,region,spp) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring")) %>% 
  mutate(spp = gsub(" ","\n",spp))

# The plot
map_plot <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_grid(spp ~ label) +
  labs(x = "Longitude", 
       y = "Latitude") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

  ggsave(filename = "proportion_chg_spp.jpg",
         plot = map_plot,
         path = my_path("R","Figures"),
         width = 14,
         height = 14
         )

@@

![Proportion change per species map for paper](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/proportion_chg_spp.jpg)

@@

state_fit <- historic_spp %>% 
  group_by(state,year,region) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  left_join(periods,
            by = "year") %>% 
  group_by(state,order,label,region) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  filter(str_detect(region,"Spring"))

agg_map <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_wrap(~ label, nrow = 1) +
  labs(x = "Longitude", 
       y = "Latitude") +
  my_ggtheme_p(facet_tx_s = 16,
               leg_pos = "bottom",
               leg_tl_s = 14,
               ax_tx_s = 10,
               ax_tl_s = 14) +
  theme(legend.key.width = unit(2,"line"))

  ggsave(filename = "proportion_chg_agg.jpg",
         plot = agg_map,
         path = my_path("R","Figures"),
         width = 9,
         height = 5
         )

@@

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
         ) %>%  
  left_join(state_lat)

# Plot me
agg_area <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  my_ggtheme_p(leg_pos = "right") +
  theme(legend.key.width = unit(1,"line"))

  ggsave(filename = "area_plot_avg_agg.jpg",
         plot = agg_area,
         path = my_path("R","Figures"),
         width = 8,
         height = 6
  )

@@

![Proportion change aggregated all species map for paper](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/area_plot_avg_agg.jpg)

@@

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region,spp) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region,spp) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
         ) %>%  
  left_join(state_lat)

# Plot me
ssp_area <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  my_ggtheme_p(leg_pos = "bottom") +
  theme(legend.key.width = unit(1,"line")) + 
  facet_wrap(~spp, ncol = 1, strip.position = "right")

  ggsave(filename = "area_plot_avg_spp.jpg",
         plot = ssp_area,
         path = my_path("R","Figures"),
         width = 10,
         height = 12
  )

```
![Proportion change aggregated all species map for paper](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/area_plot_avg_spp.jpg)

>>>>

---
title: "Spatial Analysis"
author: "Juliano Palacios Abrantes"
output:
  html_document:
  # html_notebook:
    fig_width: 6
    fig_height: 4
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float:
      collapsed: false
    theme: darkly
---

@@

# Get state order from North to south
state_order <- my_path("D","Spatial/grid_eez_fp_sf", name = "grid_eez_fp_df.csv",read = T) %>%
  group_by(state) %>% 
  summarise(
    order = mean(lat)
  ) %>% 
  mutate(abrev = c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")
  ) %>% 
  arrange(desc(order)) %>% 
  mutate(order= c("b","a",letters[3:11])) %>% # weirdly maine goes below
  arrange(order)

@@

grid_eez_fp_sf <- arrange(grid_eez_fp_sf,group)
# Plotting 
# Make sure st_simplify() is run for eez sf

@@

p_sw <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_eez_sw_sf) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color =group , fill = group), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    scale_color_manual(values = state_pallet) +
    scale_fill_manual(values = state_pallet) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45))+
    labs(x = "", y = "", title = "A) State waters approach") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot(grid_eez_sw_sf) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    geom_sf(aes(color = group),size = 0.1, alpha = 0.3) +
    facet_wrap(~ group) +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet,
                       # labels = unique(grid_eez_sw_sf$group)) +
                       labels = grid_eez_sw_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

@@

p_fp <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_eez_fp_sf) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color = group, fill = abrev), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    scale_color_manual(values = state_pallet) +
    scale_fill_manual(values = state_pallet) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45)) +
    labs(x = "", y = "", title = "B) Fishing ports approach") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot(grid_eez_fp_sf) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    geom_sf(aes(color = group),size = 0.1, alpha = 0.3) +
    facet_wrap(~ group) +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet,
                       # labels = unique(grid_eez_fp_sf$group)) +
                       labels = grid_eez_fp_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

@@

ggsave(filename = "buffer_figure.jpg",
       plot = combined_plot,
       path = my_path("R","Figures"),
       width = 11,
       height = 11
)

@@

# # single species run
run_tis(input_data = ocean_data,
        grid = grid_eez_fp_df,
        years = years,
        reg = regions,
        taxa = "Stenotomus chrysops",
        method = "fp"
)

@@

State allocations of the commercial black sea bass coastwide quota were originally implemented in 2003 as part of Amendment 13, loosely based on historical landings from 1980- 2001.

@@

spp_paths <- list.files(my_path("R",extra_path = "Partial/Interpolation/sw"))

@@

#### Load sw interpolation

taxon_sw <- my_path("R",extra_path = "Partial/Interpolation/sw", name = taxon_read)

taxon_fp <- my_path("R",extra_path = "Partial/Interpolation/fp", name = taxon_read)

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_sw, fread)) %>% 
  mutate(spatial = "sw") %>% 
  bind_rows(
    bind_rows(lapply(taxon_fp, fread))
  ) %>% 
  mutate(state = str_to_sentence(state),
         spatial = ifelse(is.na(spatial),"fp",spatial))

unique(historic_spp$spp)
unique(historic_spp$spatial)

@@

data_grid <- historic_spp %>% 
  group_by(state,lat,lon,region,spatial) %>% 
  summarise(sum = sum(value,na.rm= T), .groups = "drop") %>% # sum of all species
  group_by(state,lat,lon,region,spatial) %>% 
  summarise(mean = mean(sum,na.rm= T), .groups = "drop") # average of all years

ggplot(data = subset(data_grid, !is.na(mean))) +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = mean,
      col = mean
    )
  ) +
  geom_sf(data = land_sf, aes()) +
  facet_wrap(~ state  +  region + spatial) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_viridis("Mean Interpolation", alpha = 0.8) +
  scale_color_viridis("Mean Interpolation", alpha = 0.8) +
  theme_bw()
```

@@

total_fited <- historic_spp %>% 
  group_by(year,region,spatial) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

state_fit <- historic_spp %>% 
  group_by(state,year,region,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  left_join(periods,
            by = "year") %>% 
  group_by(state,order,label,region,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop")

@@

p <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  ggplot() +
  # geom_sf(aes(fill = mean_per)) +
  geom_sf(aes(fill = mean_per, text = paste(state, mean_per,"% of proportion"))) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_wrap(~ region + label +spatial, 
             nrow = 2) +
  labs(x = "Latitude", 
       y = "Longitude") +
  my_ggtheme_p(facet_tx_s = 12,
               leg_pos = "bottom") +
  theme(legend.key.width = unit(1,"line")) +
  theme_dark()

@@

total_fited <- historic_spp %>% 
  group_by(year,region,spp,spatial) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

state_fit <- historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100,
         label = ifelse(year >= 1980 & year <= 2001,"Reference",
                         ifelse(year > 2001,"Today",NA)
                         )
         ) %>% 
  # left_join(periods,
  #           by = "year") %>% 
  group_by(state,label,region,spp,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(label)) %>% 
  spread(spatial,mean_per) %>% 
  mutate(difference = (fp-sw)/((fp+sw)/2)*100,
         difference = ifelse(difference > 100,100,
                             ifelse(difference < -100,-100,difference)
         )
  ) %>% 
  gather("spatial","mean_per",fp:difference) %>% 
  mutate(spp = gsub(" ","\n",spp),
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State Waters","Difference")
         )
  )

# The plot
map_plot <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

diff_plot <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Percentage difference", alpha = 0.8) +
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

# Cowplot option
p <- ggdraw() +
  # Revenue circular
  draw_plot(map_plot, x = 0, y = 0.2, width = 1, height = 0.8) +
  # Catch circular
  draw_plot(diff_plot, x = 0, y = 0, width = 1, height = 0.4) +
  draw_plot_label(label = c("Latitude", "Longitude"), 
                  size = 18,
                  angle = c(90,0),
                  x = c(0,0.45), 
                  y = c(0.45,0.15)
  )

save_plot(
  path = my_path("R","Figures"),
  "proportion_chg_spp_spatial.jpg",
  p,
  base_height = 14,
  base_width = 14
)

@@

### Proportion change realtive to past

total_fited <- historic_spp %>%
  group_by(year,region,spp,spatial) %>%
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

state_change <-historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100,
         label = ifelse(year >= 1980 & year <= 2001,"Reference",
                         ifelse(year > 2001,"Today",NA)
                         )
         ) %>% 
  # left_join(periods,
            # by = "year") %>% 
  group_by(state,label,region,spp,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(label)
         ) %>% 
  select(-region) %>% 
  spread(label,mean_per) %>% 
  mutate(change = (Today-Reference)/(Today)*100,
         change = ifelse(change > 100,100,
                         ifelse(change < -100,-100,change)
         )
  ) %>%
  select(-4,-5) %>% 
  spread(spatial,change) %>% 
  mutate(difference = (fp-sw)/((fp+sw)/2)*100,
         difference = ifelse(difference > 100,100,
                             ifelse(difference < -100,-100,difference)
         )
  ) %>% 
  gather("spatial","mean_per",fp:difference) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State Waters","Difference")
         )
  )

# The plot
prop_chng_map <- land_sf %>% 
  left_join(state_change,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Change in proportion today\n relative to 1973-1984\n(%)", alpha = 0.8,
                              limits = c(-100,100),
                              breaks = seq(-100,100,25),
                              labels = c(seq(-100,75,25),">100")
                              ) +
  facet_grid(spatial~ spp) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        legend.title.align=0.5)

ggsave(filename = "temp_proportion_chng.jpg",
         plot = prop_chng_map,
         path = my_path("R","Figures"),
         width = 10,
         height = 10
  )

#### Option with differences

prop_chng_map_diff <- land_sf %>% 
  left_join(state_change,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Change in proportion\n relative to 1980-2001 (%)", 
                              alpha = 0.8,
                              # option = "C"
                              limits = c(-100,100),
                              breaks = seq(-100,100,25),
                              labels = c(seq(-100,75,25),">100")
                              ) +
  facet_grid(spatial~ spp) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        legend.title.align=0.5,
        strip.background = element_blank(),
        strip.text.x = element_blank())

# Cowplot option
p <- ggdraw() +
  # Revenue circular
  draw_plot(prop_chng_map, x = 0, y = 0.25, width = 1, height = 0.65) +
  # Catch circular
  draw_plot(prop_chng_map_diff, x = 0.101, y = 0, width = 0.798, height = 0.4) +
  draw_plot_label(label = c("Latitude", "Longitude"), 
                  size = 16,
                  angle = c(90,0),
                  x = c(0.1,0.45), 
                  y = c(0.45,0.11)
  )

save_plot(
  path = my_path("R","Figures"),
  "proportion_chg_diff.jpg",
  p,
  base_height = 14,
  base_width = 14
)

![Proportion change per species map for paper](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/temp_proportion_chng.jpg)

@@

total_fited <- historic_spp %>% 
  group_by(year,region,spatial) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

state_fit <- historic_spp %>% 
  group_by(state,year,region,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spatial")) %>%
  mutate(percentage = state_value/total_value*100, 
  # left_join(periods,
            # by = "year") %>% 
  label = ifelse(year >= 1980 & year <= 2001,"Reference",
                         ifelse(year > 2001,"Today",NA)
                         )
         ) %>% 
  group_by(state,label,region,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  filter(str_detect(region,"Spring"),
         !is.na(label)
         ) %>% 
  spread(spatial,mean_per) %>% 
  mutate(difference = (fp-sw)/((fp+sw)/2)*100,
         difference = ifelse(difference > 100,100,
                             ifelse(difference < -100,-100,difference)
         )
  ) %>% 
  gather("spatial","mean_per",fp:difference) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State Waters","Difference")
         )
  )

# Spatial map
agg_sppatial_map <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Distribution\nProportion", alpha = 0.8) +
  facet_grid(spatial~ label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 16,
               leg_pos = "right",
               leg_tl_s = 14,
               ax_tx_s = 10,
               ax_tl_s = 14) +
theme(legend.title.align=0.5)

# Difference map
agg_diff_map <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Percentage\nDifference",option = "C", alpha = 0.8) +
  facet_grid(spatial~ label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 16,
               leg_pos = "right",
               leg_tl_s = 14,
               ax_tx_s = 10,
               ax_tl_s = 14)+
  theme(
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.title.align=0.5
        )

# Cowplot option
p <- ggdraw() +
  # Revenue circular
  draw_plot(agg_sppatial_map, x = 0, y = 0.35, width = 1, height = 0.65) +
  # Catch circular
  draw_plot(agg_diff_map, x = 0.093, y = 0.05, width = 0.814, height = 0.4) +
  draw_plot_label(label = c("Latitude", "Longitude"), 
                  size = 10,
                  angle = c(90,0),
                  x = c(0,0.35), 
                  y = c(0.45,0.11)
  )

save_plot(
  path = my_path("R","Figures"),
  "proportion_chg_agg.jpg",
  p,
  base_height = 8,
  base_width = 8
)

@@

total_fited <- historic_spp %>% 
  group_by(year,region,spatial) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region,spatial) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
         ) %>%  
  left_join(state_lat) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State Waters","Difference")
         )
  )

# Plot me
agg_area <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  annotate(geom= "rect",
           xmin = 1980, xmax = 2001, 
           ymin = 0, ymax = 100,
            fill = "gray90",
           colour = NA,
           alpha = 0.5) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State\n(Latitudinal order)",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  scale_x_continuous(breaks = seq(1979,2015,9),limits = c(1979,2015))+
  xlab("Year") +
  my_ggtheme_p(leg_pos = "right", hjust = 1,ax_tx_s = 10,leg_tx_s = 10, leg_tl_s = 12,ax_tl_s = 12) +
  theme(legend.key.width = unit(1,"line"),
        legend.title.align=0.5) +
  facet_wrap(~spatial)

    ggsave(filename = "area_plot_avg_agg_sptl.jpg",
         plot = agg_area,
         path = my_path("R","Figures"),
         width = 8,
         height = 6
  )

@@

![Proportion change aggregated all species map for paper](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/area_plot_avg_agg_sptl.jpg)

@@

total_fited <- historic_spp %>% 
  group_by(year,region,spp,spatial) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region,spp,spatial) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
         ) %>%  
  left_join(state_lat) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State Waters","Difference")
         )
  )

# Plot me
ssp_area <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  annotate(geom= "rect",
           xmin = 1980, xmax = 2001, 
           ymin = 0, ymax = 100,
            fill = "gray90",
           colour = NA,
           alpha = 0.5) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State\n(Latitudinal order)",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  facet_grid(spp~spatial) +
  scale_x_continuous(breaks = seq(1979,2015,9),limits = c(1979,2015))+
  xlab("Year") +
  my_ggtheme_p(leg_pos = "right", hjust = 0.9,ax_tx_s = 14,leg_tx_s = 16, leg_tl_s = 18,ax_tl_s = 18,facet_tx_s = 16) +
  theme(legend.key.width = unit(1,"line"),
        legend.title.align=0.5)

  ggsave(filename = "area_plot_avg_spp_sptl.jpg",
         plot = ssp_area,
         path = my_path("R","Figures"),
         width = 12,
         height = 10
  )

```
![Proportion change aggregated all species map for paper](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/area_plot_avg_spp_sptl.jpg)
